### 2021-03-06 project/paper/paper.tex
tests performed

\documentclass{article}
\usepackage{pgfplotstable}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[giveninits=true]{biblatex}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{caption}
\usepgfplotslibrary{units}
\pgfplotsset{compat=1.16}

\begin{document}

\title{PDE Solvers for the Laplace's Equation}
\author{Fredrik Öberg \\ \href{mailto:fobe@kth.se}{fobe@kth.se}}
\affil{Examiner: Prof. Vladimir Vlassov}
\affil{KTH Royal Institute of Technology, Sweden}
\date{\today}
\maketitle

\begin{abstract}
Abstract
\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}
intro

\section{Programs}\label{programs}

\section{Performance Evaluation}\label{performanceevaluation}

setup

\newpage

\section{Discussion}\label{discussion}

Discussion

\section{Conclusion}\label{conclusion}

conclusion

\end{document}